Do we need detailed demographic data to forecast plant population responses to climate change?

### 2016-10-28 manuscript/manuscript.tex
final corrections before sending to MEE

Despite the weak climate effects, including climate covariates did
increase the \textcolor{blue}{mean} accuracy of forecasts for two
species: \emph{B. gracilis} and \emph{Poa secunda}
(\textcolor{blue}{Fig. S3}).
\textcolor{blue}{However, skill increases were only statistically significant at $\alpha$ = 0.05 for the QBM for \emph{B. gracilis} QBM ($t_{(279)}$ = 1.80, \emph{P} = 0.037).}
Forecast error decreased significantly with the inclusion of climate
covariates for the \emph{B. gracilis} IPM (\(t_{(280)}\) = -3.72,
\emph{P} = \textcolor{blue}{0.01}) and QBM (\(t_{(280)}\) = -3.34,
\emph{P} \textless{} 0.0001), and for the \emph{Poa secunda} IPM
(\(t_{(196)}\) = -1.90, \emph{P} \textcolor{blue}{= 0.015}) and QBM
(\(t_{(196)}\) = -2.47, \emph{P} = 0.007) (Fig. S1, Supporting
Information). In no case did including climate covariates significantly
decrease forecast skill (Table S21), despite small changes in the mean
skill (Fig. 2, \textcolor{blue}{Fig. S3}).

@@

With climate covariates included and using mean parameter values, the
accuracy of both models' forecasts declined as the distance between the
last observation and the forecast increased, but they did so at similar
rates (Fig. 4). The only exception is for \emph{Poa secunda}, where QBM
forecast accuracy remained steady as the temporal distance of the
forecast increased, whereas IPM forecast accuracy declined (Fig. 4). The
forecast horizons were short: forecast accuracy fell below \(\rho\) =
0.5 after one year for the IPM for most species, and after four years,
at most, for the QBM (Fig. 4). Across the different temporal distances
from the observation to the forecast, the IPM was never more accurate
than the QBM (\emph{P} \textgreater{} 0.05 for all one-sided
\emph{t}-tests, Table S23).
\textcolor{blue}{Likewise, the QBM was rarely more accurate the IPM, the only exceptions being for \emph{H. comata} at a temporal distance of three years ($t_{(98)}$ = 2.04, \emph{P} = 0.032) and \emph{B. gracilis} at a temporal distance of eight years ($t_{(37)}$ = 1.67, \emph{P} = 0.05) (Table S24).}
\st{There were some cases where the QBM was more accurate than the IPM for \emph{Poa secunda}, but neither model exceeded the forecast proficiency threshold by a large margin (Fig. 4, Table S24).}

@@

This work was funded by the National Science Foundation through a
Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370), award
MSB-1241856 to MBH, and a CAREER award to PBA (DEB-1054040). We thank
the original mappers of the permanent quadrats in Montana and the
digitizers in the Adler lab, without whom this work would not have been
possible. Informal conversations with Stephen Ellner, Giles Hooker,
Robin Snyder, and a series of meetings between the Adler and Weecology
labs at USU sharpened our thinking. Brittany Teller provided comments
that improved our manuscript. Compute, storage and other resources from
the Division of Research Computing in the Office of Research and
Graduate Studies at Utah State University are gratefully acknowledged.
Any use of trade, firm, or product names is for descriptive purposes
only and does not imply endorsement by the U.S. government. This
research was supported by the Utah Agricultural Experiment Station, Utah
State University, and approved as journal paper number 8917.
\textcolor{blue}{And, thanks to Andy Kleinhesselink for discovering an error in our computer code before this paper was published.}

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[height=6in]{./components/forecast_accuracy_mockup_boxplot.png}
  \caption{Comparisons of one-step-ahead, out-of-sample forecast accuracy between the IPM and QBM models with and without the inclusion of climate covariates. Boxplots show the distribution of $\rho$ averaged over quadrats for each cross-validation year (i.e., 13 values of $\rho$ for each species-model combination). For each comparison, \emph{P}-values are from one-sided \emph{t} tests designed to assess whether the first model in the comparison statement had higher accuracy than the second model in the comparison statement (see details in Table S22). Statistical tests relied on correlation values for each quadrat-year-species combination, after averaging over model reps for each combination. In no case did adding climate covariates decrease forecast accuracy (Table S21). Species codes are as in Fig. 1.}
\end{figure}

>>>>

Despite the weak climate effects, including climate covariates did
increase the \textcolor{blue}{mean} accuracy of forecasts for two
species: \emph{B. gracilis} and \emph{Poa secunda} (Fig.2 and
\textcolor{blue}{Fig. S3, Supporting Information}).
\textcolor{blue}{However, skill increases were only statistically significant at $\alpha$ = 0.05 for the \emph{B. gracilis} QBM ($t_{(279)}$ = 1.80, \emph{P} = 0.037).}
Forecast error decreased significantly with the inclusion of climate
covariates for the \emph{B. gracilis} IPM (\(t_{(280)}\) = -3.72,
\emph{P} = \textcolor{blue}{0.01}) and QBM (\(t_{(280)}\) = -3.34,
\emph{P} \textless{} 0.0001), and for the \emph{Poa secunda} IPM
(\(t_{(196)}\) = -1.90, \emph{P} \textcolor{blue}{= 0.015}) and QBM
(\(t_{(196)}\) = -2.47, \emph{P} = 0.007) (Fig. S1, Supporting
Information). In no case did including climate covariates significantly
decrease forecast skill (Table S21), despite small changes in the mean
skill (Fig. 2, \textcolor{blue}{Fig. S3}).

@@

With climate covariates included and using mean parameter values, the
accuracy of both models' forecasts declined as the distance between the
last observation and the forecast increased, but they did so at similar
rates (Fig. 4). The only exception is for \emph{Poa secunda}, where QBM
forecast accuracy remained steady as the temporal distance of the
forecast increased, whereas IPM forecast accuracy declined (Fig. 4). The
forecast horizons were short: forecast accuracy fell below \(\rho\) =
0.5 after one year for the IPM for most species, and after four years,
at most, for the QBM (Fig. 4). Across the different temporal distances
from the observation to the forecast, the IPM was never more accurate
than the QBM (\emph{P} \textgreater{} 0.05 for all one-sided
\emph{t}-tests, Table S23).
\textcolor{blue}{Likewise, the QBM was rarely more accurate than the IPM, the only exceptions being for \emph{H. comata} at a temporal distance of three years ($t_{(98)}$ = 2.04, \emph{P} = 0.032) and \emph{B. gracilis} at a temporal distance of eight years ($t_{(37)}$ = 1.67, \emph{P} = 0.05) (Table S24).}
\st{There were some cases where the QBM was more accurate than the IPM for \emph{Poa secunda}, but neither model exceeded the forecast proficiency threshold by a large margin (Fig. 4, Table S24).}

@@

This work was funded by the National Science Foundation through a
Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370), award
MSB-1241856 to MBH, and a CAREER award to PBA (DEB-1054040). We thank
the original mappers of the permanent quadrats in Montana and the
digitizers in the Adler lab, without whom this work would not have been
possible. Informal conversations with Stephen Ellner, Giles Hooker,
Robin Snyder, and a series of meetings between the Adler and Weecology
labs at USU sharpened our thinking. Brittany Teller provided comments
that improved our manuscript. Compute, storage and other resources from
the Division of Research Computing in the Office of Research and
Graduate Studies at Utah State University are gratefully acknowledged.
Any use of trade, firm, or product names is for descriptive purposes
only and does not imply endorsement by the U.S. government. This
research was supported by the Utah Agricultural Experiment Station, Utah
State University, and approved as journal paper number 8917.

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[height=6in]{./components/forecast_accuracy_mockup_boxplot.png}
  \caption{Comparisons of one-step-ahead, out-of-sample forecast accuracy between the IPM and QBM models with and without the inclusion of climate covariates. Boxplots show the distribution of $\rho$ averaged over quadrats for each cross-validation year (i.e., 13 values of $\rho$ for each species-model combination). For each comparison, \emph{P}-values are from one-sided \emph{t} tests designed to assess whether the first model in the comparison statement had higher accuracy than the second model in the comparison statement (see details in Table S22). \textcolor{blue}{A single asterisk indicates a statistically significant difference at $\alpha$ = 0.1 and double asterisks indicate significance at $\alpha$ = 0.05}. Statistical tests relied on correlation values for each quadrat-year-species combination, after averaging over model reps for each combination. In no case did adding climate covariates decrease forecast accuracy (Table S21). Species codes are as in Fig. 1.}
\end{figure}
